Machine Learning Methods in Functional Programming and Interactive Theorem Proving

### 2015-10-30 TransferReport/report.tex
Imported feature extraction writeup

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
%\usepackage{enumitem}
\usepackage{paralist}
\usepackage{csquotes}
\usepackage[affil-it]{authblk}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,decorations.pathreplacing,calc}

@@

Haskell is a programming language which is well-suited to research; indeed, this was a goal of the language's creators \cite{marlow2010haskell}. Like most members of the \emph{functional programming} paradigm, Haskell is essentially a variant of $\lambda$-calculus; in fact the \textsc{GHC} compiler uses (variants of) the polymorphic $\lambda$-calculus \emph{System F} as an intermediate representation (referred to a \emph{GHC Core}). For a detailed exposition of the language syntax and semantics, you can consult the Haskell Language Report \cite{marlow2010haskell} \footnote{Version 2010 was current as this was written}; here we will focus on those language features which make it particularly suited for our formal reasoning purposes:

@@

\subsection{Intention}

One consideration when generating conjectures is the difference (if any) between theorems, lemmas, corollaries, etc. From a logical point of view, these are all equivalent; it Most proof assistants do not distinguish between these, although some (like Coq) may allow users to annotate their statements with label.

\subsection{QuickCheck Properties}

@@

% TODO: Examples
% TODO: \cite{Schmidhuber1999} Bet

@@

% TODO: \cite{Kaplan2006}
% TODO: \cite{Lipson2007}
% TODO: \cite{Luciw2011}
% TODO: \cite{Macedo2000}
% TODO: \cite{Ramik.Sabourin.Madani:2013}
% TODO: \cite{Roa.Kruijff.Jacobsson:2009}
% TODO: \cite{Schaul.Sun.Wierstra.ea:2011}
% TODO: \cite{Schmidhuber1999}
% TODO: \cite{Schmidhuber:1991}
% TODO: \cite{Scott1989}
% TODO: \cite{Steunebrink.Koutnik.Thorisson.ea:2013}
% TODO: \cite{maher2008achieving}
% TODO: \cite{meyer1991possibility}
% TODO: \cite{oudeyer2004intelligent}
% TODO: \cite{oudeyer2014evolution}
% TODO: \cite{schmidhuber2006developmental}

@@

The combinatorial nature of formal systems leads many algorithms in this domain, such as proof search methods, to have exponential time complexityCITE; hence even a modest size increase can turn a trivial problem into an intractable one. It is currently infeasible to improve the scaling behaviour of many such algorithms; for example, problems like TODO are in the infamous NP-complete complexity classCITE. On the other hand, we can turn this difficulty around: a modest \emph{decrease} in size may turn an intractable problem into a solvable one. We can ensure that the solutions to these reduced problems coincide with the original if we only remove \emph{redundant} information. This leads to the idea of \emph{relevance filtering}.

Relevance filtering simplifies a proof search problem by removing from consideration those axioms, definitions, lemmas, etc. which are deemed \emph{irrelevant}. The technique is used in Sledgehammer during its translation of Isabelle/HOL theories to statements in first order logic: rather than translating the entire theory, only those lemmas which are deemed ``relevant'' (required for proving the statement) are included. This reduces the size of the problem and speeds up the proof search, but it creates the new problem of determining when a lemma is relevant: how do we know what will be required, before we have the proof?

Sledgehammer's approach is to use \emph{supervised learning} to approximate the relevance relation: a naive bayes classifier is trained on TODO

ML4PG
ACL2(ml)
Josef Urban
Sledgehammer
Naive Bayes

@@

The most important property of a logical formula is its truth value. Although we may be able to determine some truth values exactly, eg. using decision or semi-decision procedures, it may be more efficient to \emph{approximate} truth values. One straightforward extension of truth values is \emph{probabilities}, where we can assign probability $1$ to formulae which are known to be true, $0$ to formulae known to be false, and intermediate values to those which we do not yet know. This provides a   may be able to determine this truth value, it may not always be even an interesting formula
TODO: \cite{Hutter.Lloyd.Ng.ea:2013}

@@

TODO: \cite{Montano-Rivas.McCasland.Dixon.ea:2012}
TODO: \cite{Piantadosi.Tenenbaum.Goodman:2012}
TODO: \cite{Wille:2005}
TODO: \cite{colton1999automatic}
TODO: \cite{colton2000agent}
TODO: \cite{colton2012automated}
TODO: \cite{lenat1977automated}
TODO: \cite{mullerunderstanding}
TODO: \cite{Bundy.Cavallo.Dixon.ea:2015}
TODO: \cite{johansson2009isacosy}
TODO: \cite{spector2008genetic}
TODO: https en.wikipedia.org/wiki/Discovery system

However, this search space grows exponentially in the length of the proofs, which is unfortunate since proof length has been proposed as an approximate measure of how interesting a theorem is \cite[\S~10.2.1]{colton2012automated}.

@@

% Genetic programming? GP in finite algebras

@@

The application of \emph{kernel methods} to structured information is discussed in \cite{Gartner2003}, where the input data (including sequences, trees and graphs) are represented using \emph{generative models}, such as hidden Markov models, and it is these models which are
operate alloe If we for example, a statement like may from the nested form of logical  example, logical terms may be nested to arbitrary depths a large amount of information is represented   to a domain like formal mathematics is the question of how to deal with structure.

TODO: \cite{Gartner2003}
TODO: \cite{Oveisi.Oveisi.Erfanian.ea:2012}
TODO: \cite{bakir2007predicting}
TODO: \cite{conf/ijcai/Plate91}
TODO: \cite{goller1996learning}
TODO: \cite{kwasny1995tail}
TODO: \cite{pollack1990recursive}
TODO: \cite{zanzotto2012distributed}

@@

One way to apply fixed-size machine learning algorithms to arbitrarily-sized recursive structures is to use a \emph{distributed representation}. These mix information from all parts of a structure together into a fixed number of bits, storing an \emph{approximation} whose accuracy depends on the size of the value and the amount of storage used.

>>>>

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
%\usepackage{enumitem}
\usepackage{paralist}
\usepackage{csquotes}
\usepackage[affil-it]{authblk}
\usepackage{listings}

% Treat paragraph as subsubsubsection
\usepackage{titlesec}
\setcounter{secnumdepth}{4}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,decorations.pathreplacing,calc}

\DeclareMathOperator{\md5}{MD5}

@@

Haskell is a programming language which is well-suited to research; indeed, this was a goal of the language's creators \cite{marlow2010haskell}. Like most members of the \emph{functional programming} paradigm, Haskell is essentially a variant of $\lambda$-calculus; in fact the \textsc{GHC} compiler uses (variants of) the polymorphic $\lambda$-calculus \emph{System F} as an intermediate representation (referred to as \emph{GHC Core}). For a detailed exposition of the language syntax and semantics, you can consult the Haskell Language Report \cite{marlow2010haskell} \footnote{Version 2010 was current as this was written} or a textbook such as \cite{thompson1999haskell}; here we will focus on those language features which make it particularly suited for our formal reasoning purposes:

@@

\subsection{Lemmas}

One consideration when generating conjectures is the difference between theorems, lemmas, corollaries, etc. From a logical point of view, these are all equivalent, and hence most proof assistants do not distinguish between them. However, their \emph{intention} may be different, and hence some systems, like Coq, allow users to \emph{label} each statement as being a \coq{Theorem}, a \coq{Lemma}, etc. despite their internal representations being the same.

One striking observation is that lemmas outnumber theorems; in the Coq standard library there are over five times as many lemmas as theorems \footnote{The latest version as of writing is \texttt{coq-8.4pl6} which, when excluding comments, includes 1492 occurences of \coq{Theorem} and 7594 of \coq{Lemma} in its \texttt{theories/} directory.}

\subsection{Tests}

Since we are working in the domain of Haskell programs, an abundant source of statements are available in the form of \emph{tests}. For manually-written tests, the effort required to write them implies that they must be of some interest to their author. Whilst many forms of test are \emph{not} suitable for our purposes, such as \emph{unit tests} which are always trivially provable by $\beta$-reduction; or \emph{integration tests}, which depend on the behaviour of the external environment. One reason to study Haskell is its widespread use of \emph{property checking}, which does give us useful data. Many Haskell property checkers exist, based on random testing (\textsc{QuickCheck} \cite{claessen2011quickcheck} and \textsc{SmartCheck} \cite{pike2014smartcheck}), enumeration (\textsc{SmallCheck} \cite{runciman2008smallcheck}), observation (\textsc{Lazy SmallCheck} \cite{reich2013advances}) and logic programming (\textsc{SparseCheck} \cite{sparsecheck}).

Thankfully the major differences between these systems are in the way they instantiate test arguments; their representations of properties are largely the same (modulo renaming of functions and types).

Here we consider 10 most-downloaded packages from hackage.haskell.org (as of 2015-10-30) which have property tests; these are \texttt{warp}, \texttt{aeson}, \texttt{text}, \texttt{lens}, \texttt{conduit}, \texttt{pandoc}, \texttt{attoparsec}, \texttt{scientific}, \texttt{yesod-core} and \texttt{blaze-html}.

@@

% TODO: Examples
\cite{Schmidhuber1999}

@@

% TODO
\cite{Kaplan2006}
\cite{Lipson2007}
\cite{Luciw2011}
\cite{Macedo2000}
\cite{Ramik.Sabourin.Madani:2013}
\cite{Roa.Kruijff.Jacobsson:2009}
\cite{Schaul.Sun.Wierstra.ea:2011}
\cite{Schmidhuber1999}
\cite{Schmidhuber:1991}
\cite{Scott1989}
\cite{Steunebrink.Koutnik.Thorisson.ea:2013}
\cite{maher2008achieving}
\cite{meyer1991possibility}
\cite{oudeyer2004intelligent}
\cite{oudeyer2014evolution}
\cite{schmidhuber2006developmental}

@@

The combinatorial nature of formal systems causes many proof search methods such as resolution, to have exponential complexity \cite{haken1985intractability}; hence even a modest size increase can turn a trivial problem into an intractable one. Finding efficient alternatives for such algorithms, especially those which are NP-complete (e.g. determining satisfiability) or co-NP-complete (e.g. determining tautologies), seems unlikely, as it would imply progress on the famously intractable open problems of $\text{P} = \text{NP}$ and $\text{NP} = \text{co-NP}$. On the other hand, we can turn this difficulty around: a modest \emph{decrease} in size may turn an intractable problem into a solvable one. We can ensure that the solutions to these reduced problems coincide with the original if we only remove \emph{redundant} information. This leads to the idea of \emph{relevance filtering}.

Relevance filtering simplifies a proof search problem by removing from consideration those clauses (axioms, definitions, lemmas, etc.) which are deemed \emph{irrelevant}. The technique is used in Sledgehammer during its translation of Isabelle/HOL theories to statements in first order logic: rather than translating the entire theory, only a sub-set of relevant clauses are included. This reduces the size of the problem and speeds up the proof search, but it creates the new problem of determining when a clause is relevant: how do we know what will be required, before we have the proof?

The initial approach, known as \textsc{MePO} (from \emph{Meng-Paulson} \cite{meng2009lightweight}), gives each clause a score based on the proportion $m / n$ of its symbols which are ``relevant'' (where $n$ is the number of symbols in the clause and $m$ is the number which are relevant). Initially, the relevant symbols are those which occur in the goal, but whenever a clause is found which scores more than a particular threshold, all of its symbols are then also considered relevant. There are other heuristics applied too, such as increasing the score of user-provided facts (e.g. given by keywords like \texttt{using}), locally-scoped facts, first-order facts and rarely-occuring facts. To choose $r$ relevant clauses for an ATP invocation, we simply order the clauses by decreasing score and take the first $r$ of them.

Recently, a variety of alternative algorithms have also been investigated, including:

  \item{\textsc{MaSH}}: Machine Learning for SledgeHammer \cite{kuhlwein2013mash}. The distinguishing feature of \textsc{MaSH} is its use of ``visibility'', which is essentially a dependency graph of which theorems were used in the proofs of which other theorems; although theorems are represented as abstract sets of features. To select relevant clauses for a goal, the set of clauses which are visible from the goal's components is generated; this is further reduced by (an efficient approximation of) a naive Bayes algorithm.

  \item{\textsc{MOR}}: \emph{Multi-output ranking} uses a support vector machine (SVM) approach for selecting relevant axioms from the Mizar Mathamatical Library for use by the Vampire ATP system \cite{alama2014premise}. \iffalse TODO: describe the kernel, as that's the interesting bit \fi It compares favourably to \textsc{SNoW} and \textsc{SInE}.

  % TODO:
  \item{\textsc{SInE}}
  \item{\textsc{BliStr}}
  \item{\textsc{HOLyHammer}}
  \item{\textsc{MoMM}}
  \item{\textsc{SNoW}}
  \item{\textsc{MPTP 0.2}}
  \item{\textsc{MaLARea}}
  \item{\textsc{MaLARea SG1}}

\subsubsection{Clustering}
% TODO: ML4PG
% TODO: ACL2(ml)

@@

The most important property of a logical formula is its truth value. Although we may be able to determine some truth values exactly, eg. using decision or semi-decision procedures, it may be more efficient to \emph{approximate} truth values. One straightforward extension of truth values is \emph{probabilities}, where we can assign probability $1$ to formulae which are known to be true, $0$ to formulae known to be false, and intermediate values to those which we do not yet know.

% TODO
\cite{Hutter.Lloyd.Ng.ea:2013}

@@

% TODO:
\cite{Montano-Rivas.McCasland.Dixon.ea:2012}
\cite{Piantadosi.Tenenbaum.Goodman:2012}
\cite{Wille:2005}
\cite{colton1999automatic}
\cite{colton2000agent}
\cite{colton2012automated}
\cite{lenat1977automated}
\cite{mullerunderstanding}
\cite{Bundy.Cavallo.Dixon.ea:2015}
\cite{johansson2009isacosy}
\cite{spector2008genetic}
\cite{colton2012automated}
% TODO: How does https en.wikipedia.org/wiki/Discovery system relate?

% However, this search space grows exponentially in the length of the proofs, which is unfortunate since proof length has been proposed as an approximate measure of how interesting a theorem is \cite[\S~10.2.1]{colton2012automated}.

@@

@@

\paragraph{Feature Extraction}

\emph{Feature extraction} is a common pre-processing step for machine learning (ML). Rather than feeding "raw" data straight into our ML algorithm, we only learn a sample of \emph{relevant} details, known as \emph{features}. This has two benefits:

\begin{itemize}
  \item \emph{Feature vectors} (ordered sets of features) are chosen to be more compact than the data they're extracted from: feature extraction is \emph{lossy compression}. This reduces the size of the ML problem, improving efficiency (eg. running time).
  \item We avoid learning irrelevant details such as the encoding system used, improving \emph{data} efficiency (the number of samples required to spot a pattern).
\end{itemize}

Another benefit of feature extraction is to \emph{normalise} the input data to a fixed-size representation. Many ML algorithms only work with fixed-size inputs; for example, the popular \emph{backpropagation} \cite{Russell:2003:AIM:773294} algorithm works on models with \emph{fixed} topology (eg. \emph{artificial neural networks} with fixed connections between nodes). This requires some form of pre-processing in domains where the size of each input is not known, may vary or may even be unbounded.

For example, in the case of \emph{online} learning we must make predictions/decisions before seeing all of the inputs. Unbounded input appears in domains such as programming and theorem proving, where individual term may be trees of unbounded depth. In these situations we need a mapping from arbitrary inputs to a fixed representation which is amenable to learning.

\paragraph{Significance}

As an example, say we want to learn relationships between the following program fragments:

\begin{lstlisting}[language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
data Maybe a = Nothing | Just a

data Either a b = Left a | Right b
\end{lstlisting}

We might hope our algorithm discovers relationships like:

\begin{itemize}
  \item Both are valid Haskell code.
  \item Both describe datatypes.
  \item Both datatypes have two constructors.
  \item \hs{Either} is a generalisation of \hs{Maybe} (we can define \hs{Maybe a = Either () a} and \hs{Nothing = Left ()}).
  \item There is a symmetry in \hs{Either}: \hs{Either a b} is equivalent to \hs{Either b a} if we swap occurences of \hs{Left} and \hs{Right}.
  \item It is trivial to satisfy \hs{Maybe} (using \hs{Nothing}).
  \item It is not trivial to satisfy \hs{Either}; we require an \hs{a} or a \hs{b}.
\end{itemize}

However, this is too optimistic. Without our domain-knowledge of Haskell, an ML algorithm cannot impose any structure on these fragments, and will treat them as strings of bits. Our high-level hopes are obscured by low-level details: the desirable patterns of Haskell types are mixed with undesirable patterns of ASCII bytes, of letter frequency in English words, and so on.

\paragraph{Practical Considerations}

In theory we could throw more computing resources and data at a problem, but available hardware and corpora are always limited. Instead, feature extraction lets us narrow the ML problem to what we, with our domain knowlege, consider important.

There is no \emph{fundamental} difference between raw representations and features: the identity function is a valid feature extractor. Likewise, there is no crisp distinction between feature extraction and machine learning: a sufficiently-powerful learner doesn't require feature extraction, and a sufficiently-powerful feature extractor doesn't require any learning! \footnote{Consider a classification problem, to assign a label $l \in L$ to each input. If we only extract a single feature $f \in L$, we have solved the classification problem without using a separate learning step.}

Rather, the terms are distinguished for purely \emph{practical} reasons: by separating feature extraction from learning, we can distinguish straightforward, fast data transformation (feature extraction) from complex, slow statistical analysis (learning). This allows for modularity, separation of concerns, and in particular allows ``off-the-shelf'' ML to be re-used across a variety of different domains.

Even if we have no domain knowledge, we can still use a feature extraction phase to improve efficiency: first we learn a compact representation for our data, for example using \emph{autoencoding}; then we use that encoder as a feature extractor for our main learning task. This stacking of one learning algorithm on top of another, especially with greedy learning of each layer, has lead to the recent trend of \emph{deep learning}.

\paragraph{Truncation and Padding}

The simplest way to limit the size of our inputs is to truncate anything larger than a particular size (and pad anything smaller). This is the approach taken by ML4PG \cite{journals/corr/abs-1302-6421}, which limits itself to trees with at most 10 levels and 10 elements per level; each tree is converted to a $30 \times 10$ matrix (3 values per tree node) and learning takes place on these normalised representations.

Truncation is unsatisfactory in the way it balances \emph{data} efficiency with \emph{time} efficiency. Specifically, truncation works best when the input data contains no redundancy and is arranged with the most significant data first (in a sense, it is ``big-endian''). The less these assumptions hold, the less we can truncate. Since many ML algorithms scale poorly with input size, we would prefer to eliminate the redundancy using a more aggressive algorithm, to keep the resulting feature size as low as possible.

\paragraph{Dimension Reduction}

A more sophisticated approach to the problem of reducing input size is to view it as a \emph{dimension reduction} technique: our inputs can be modelled as points in high-dimensional spaces, which we want to project into a lower-dimensional space ($\left\{ {0, 1} \right\}^N$ in the case of $N$-bit vectors).

Truncation is a trivial dimension reduction technique: take the first $N$ coordinates (bits). More sophisticated projection functions consider the \emph{distribution} of the points, and project with the hyperplane which preserves as much of the variance as possible (or, equivalently, reduces the \emph{mutual information} between the points).

There are many techniques to find these hyperplanes, such as \emph{principle component analysis} (PCA) and \emph{autoencoding}; however, since these techniques are effectively ML algorithms in their own right, they suffer some of the same constraints we're trying to avoid:

\begin{itemize}
  \item They operate *offline*, requiring all input points up-front
  \item All input points must have the same dimensionality
\end{itemize}

In particular, the second constraint is precisely what we're trying to avoid. Sophisticated dimension reduction is still useful for \emph{compressing} large, redundant features into smaller, information-dense representations, and as such provides a good complement to truncation.

The requirement for offline ``batch'' processing is more difficult to overcome, since any learning we perform for feature extraction will interfere with the core learning algorithm that's consuming these features. Such systems are possible, but out of scope for this small initial implementation.

\paragraph{Structured Data}

The task of dimension reduction changes when we consider \emph{structured} data. Recursive structures, like trees and lists, have \emph{fractal} dimension: adding layers to a recursive structure gives us more \emph{fine-grained} features, rather than \emph{orthogonal} features. For data mining context-free languages (eg. those of programming and theorem-proving systems), we will mainly be concerned with tree structures of the following form (where \hs{FV} denotes a feature vector):

\begin{lstlisting}[language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
data Tree = Leaf FV
          | Branch FV Tree Tree
\end{lstlisting}

\paragraph{Truncation}

Just like the non-recursive case, the simplest way to reduce our dimensionality is to truncate. To reduce an arbitrary \hs{t :: Tree} to a maximum depth \hs{d > 0}, we can use \hs{reduceD d t} where:

\begin{lstlisting}[language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
reduceD 1 (Branch fv l r) = Leaf   fv
reduceD n (Leaf   fv)     = Leaf   fv
reduceD n (Branch fv l r) = Branch fv (reduceD (n-1) l)
                                      (reduceD (n-1) r)
\end{lstlisting}

This is simple, but suffers two problems:

\begin{itemize}
 \item We must choose a conservatively large \hs{d}, since we're throwing away information
 \item The number of feature vectors grows exponentially as \hs{d} increases
\end{itemize}

The first problem can't be avoided when truncating, but the second can be mitigated by truncating the \emph{width} of the tree as well, to some \hs{w > 0}. One way to truncate the width is tabulating the tree's feature vectors, then truncating the table to \hs{w} $\times$ \hs{d}. We can do this with \hs{tabulate w d t}:

\begin{lstlisting}[language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
tabulate w d t = take d (map (take w) (tabulate' t))

tabulate' (Leaf   fv)     = [fv] : []
tabulate' (Branch fv l r) = [fv] : merge (tabulate' l) (tabulate' r)

merge    []     ys  = ys
merge    xs     []  = xs
merge (x:xs) (y:ys) = (x ++ y) : merge xs ys
\end{lstlisting}

\paragraph{Extracting Features from Haskell}

The Glasgow Haskell Compiler provides a library interface to its operation, which we can use to parse Haskell code into tree structures. It also provides a \emph{renaming} transformation, which makes all names unique; either by prefixing them with the name of the module which defines them, or by suffixing the name with a number.

\paragraph{Linear Combinations}

Rather than truncating our trees, we can \emph{combine} the feature vectors of leaves into their parents, and hence \emph{fold} the tree up to any finite depth. Following \cite{zanzotto2012distributed} we can use \emph{circular convolution} (\hs{cc}) as our combining function. Hence, to reduce \hs{t :: Tree} to a single feature vector, we can use \hs{reduceC t} where:

\begin{lstlisting}[language=Haskell, xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
reduceC (Leaf fv)       = fv
reduceC (Branch fv l r) = cc fv (cc (reduceC l) (reduceC r))
\end{lstlisting}

Circular convolution has the following desirable properties:

  \item{Non-commutative}: \hs{cc a b} $\not\approx$ \hs{cc b a}, hence we can distinguish between \hs{Branch fv a b} and \hs{Branch fv b a}

  \item{Non-associative}: \hs{cc a (cc b c)} $\not\approx$ \hs{cc (cc a b) c}, hence we can distinguish between \hs{Branch v1 a (Branch v2 b c)} and \hs{Branch v1 (Branch v2 a b) c)}.

Here we use $\not\approx$ to represent that these quantities differ \emph{with high probability}, based on empirical testing. This is the best we can do, since the type of combining functions \hs{(FV, FV) -> FV} has a co-domain strictly smaller than its domain, and hence it must discard some information. In the case of circular convolution, the remaining information is spread out among the bits in the resulting feature, which is known as a \emph{distributed representation} \cite{plate1991holographic}.

\paragraph{Extracting Features from XML}

This method of folding with circular convolution has been implemented in the Haskell program \textsc{Tree Features} \footnote{Available online at \href{http://chriswarbo.net/git/tree-features}{http://chriswarbo.net/tree-features}. The application parses arbitrary trees of XML and uses binary features, ie. feature vectors are bit vectors.

To calculate the initial feature of an XML element, before any folding occurs, we concatenate its name and attributes to produce a string \hss}. The feature is then:

\begin{equation}
feature(n) = 2^{\md5(s) \pmod{L}}
\end{equation}

Where $\md5$ is the MD5 hash algorithm and $L$ is the length of the desired vector, given as a commandline argument. The result of $feature(s)$ is a bit vector of length $L$, containing only a single $1$. Due to the use of a hashing function, the position of that $1$ can be deterministically calculated from the input $s$, yet the \emph{a priori} probability of each position is effectively uniform.

In other words, the hash introduces unwanted patterns which are \emph{theoretically} learnable, but in practice a strong hash can be treated as irreversible and hence unlearnable.

\paragraph{Application to Coq}

Prior to 2014-09-08, the Coq proof assistant provided a rudimentary XML import/export feature, which we can use to access tree structures for learning. We achieve this by using the \coq{external} primitive of the Ltac tactic language: \coq{external "treefeatures" "32" t1 t2 t3.} will generate an XML tree representing the Coq terms \coq{t1}, \coq{t2} and \coq{t3}, and send it to the standard input of a \texttt{treefeatures} invocation, using a vector length argument of 32.

Coq will also interpret the standard output of the command, to generate terms and tactics. This functionality isn't yet used by \textsc{Tree Features}, but it is clear that a feedback loop can be constructed, allowing for powerful ML approaches like \emph{reinforcement learning}.

\paragraph{Sequencing}

Any investigation of variable-size input would be incomplete without mentioning the \emph{sequencing}. This is a lossless approach, which splits the input into fixed-size \emph{chunks}, which are fed into an appropriate ML algorithm one at a time. The sequence is terminated by a sentinel; an ``end-of-sequence'' marker which, by construction, is distinguishable from the data chunks. This technique allows us to trade \emph{space} (the size of our input) for \emph{time} (the number of chunks in a sequence).

Not all ML algorithms can be adapted to accept sequences. One notable approach is to use \emph{recurrent ANNs} (RANNs), which allow arbitrary connections between nodes, including cycles. Compared to \emph{feed-forward} ANNs (FFANNs), which are acyclic, the \emph{future output} of a RANN may depend arbitrarily on its \emph{past inputs} (in fact, RANNs are universal computers).

The main problem with RANNs, compared to the more widely-used FFANNs, is the difficulty of training them. If we extend the standard backpropagation algorithm to handle cycles, we get the \emph{backpropagation through time} algorithm \cite{werbos1990backpropagation}. However, this suffers a problem known as the \emph{vanishing gradient}: error values decay exponentially as they propagate back through the cycles, which prevents effective learning of delayed dependencies, undermining the main advantage of RANNs. The vanishing gradient problem is the subject of current research, with countermeasures including \emph{neuroevolution} (using evolutionary computation techniques to train an ANN) and \emph{long short-term memory} (LSTM; introducing a few special, untrainable nodes to persist values for long time periods \cite{hochreiter1997long}).

The application of \emph{kernel methods} to structured information is discussed in \cite{Gartner2003}, where the input data (including sequences, trees and graphs) are represented using \emph{generative models}, such as hidden Markov models, of a fixed size.

% TODO
\cite{Gartner2003}
\cite{Oveisi.Oveisi.Erfanian.ea:2012}
\cite{bakir2007predicting}
\cite{conf/ijcai/Plate91}
\cite{goller1996learning}
\cite{kwasny1995tail}
\cite{pollack1990recursive}
\cite{zanzotto2012distributed}

\iffalse

@@

One way to apply fixed-size machine learning algorithms to arbitrarily-sized recursive structures is to use a \emph{distributed representation}. These mix information from all parts of a structure together into a fixed number of bits, storing an \emph{approximation} whose accuracy depends on the size of the value and the amount of storage used.
\fi
